\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{lipsum} % for dummy text if needed
\usepackage{xcolor}
\geometry{margin=1in}

\usepackage[backend=biber, style=authoryear]{biblatex}

\addbibresource{refs.bib} % BibTeX file

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\prox}{\operatorname{prox}}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{assumption}{Assumption}

\setlength{\parindent}{0pt}

\title{Project Proposal on Qualitative Analysis of Machine Unlearning Algorithms}
\author{Kevin K Thomas \\ Inzaghi Moniaga}
\date{}

\begin{document}
\maketitle

\section{Introduction}
In the era of big data, the ability to unlearn specific data from trained models has become increasingly important for privacy reasons. 
Although retraining a model from scratch to forget data is the best approach, it is computationally expensive. This motivates the need for efficient approximate unlearning algorithms that do not require retraining. Following the 2023 NeurIPS Unlearning Competition \cite{neurips-2023-machine-unlearning}, several new approaches have been proposed. 
This project compares their qualitative performance against the 2023 winning solution (\cite{2023-winner}) and aims to provide a survey of the different approaches.

\section{Problem setup}
We base our problem setup on \cite{li2025overview}. We begin with some notation and definitions. Let $Z$ be the set of all data points and $Z^*$ be the set of all possible training datasets. Let $H$ be the set of all possible hypotheses (i.e. parameters). We define
\begin{definition}[Learning Algorithm]
    A learning algorithm $A: Z^* \to H$ maps a training dataset to a hypothesis. If $D \in Z^*$, then $A(D)$ is the model obtained by training on $D$.
\end{definition}
\begin{definition}[Unlearning Algorithm]
    An unlearning algorithm $U: H \times Z^* \times Z^* \to H$ takes a trained model, the original training dataset, and a subset of data points to unlearn, and produces a new model. Specifically, if $h = A(D)$ is the model trained on $D$, and $S \subseteq D$ is the subset of data points to unlearn, then $U(h, D, S)$ is the model obtained by unlearning $S$ from $h$.
\end{definition}
Note that both $A$ and $U$ can be randomized algorithms that define a distribution over the hypothesis space $H$. \\

Suppose we have a training dataset $D \in Z^*$ and a subset of data points $S \subseteq D$ that we wish to unlearn. Let $D' = D \setminus S$. The goal of machine unlearning is to create an unlearning algorithm $U$ such that the distribution of $U(A(D), D, S)$ is close to the distribution of $A(D')$ (i.e. retraining the original model on the smaller dataset).
i.e.
\begin{equation}
    \Pr[A(D')] \approx \Pr[U(A(D), D, S)]
\end{equation} 
Of course, we do not have access to the true distributions so there are several metrics that have been proposed to evaluate the performance of unlearning algorithms. One common metric is based on differential privacy. (See \cite{dwork2014algorithmic} and \cite{sekhari2021remember}). The NeurIPS 2023 Unlearning Competition uses metrics based on differential privacy. \\

\section{What do you plan to do?}
We plan to do a comparative review, benchmark and survey of more recent unlearning algorithms for image classification. We will use ResNet-18 as a baseline model, retrain from scratch to get an exact unlearning baseline, and evaluate the same metrics used for the Kaggle competition (\cite{neurips-2023-machine-unlearning}).

The algorithm we currently plan to compare are:
\begin{itemize}
    \item The 2023 NeurIPS Unlearning Competition winning solution [\cite{2023-winner}]
    \item Masked Small Gradient with Convolution Transpose (CT) [\cite{cadet2025}]
    \item SCRUB [\cite{kurmanji2023unboundedmachineunlearning}]
    \item DEcoupLEd Distillation To Erase (DELETE) [\cite{zhou2025decoupleddistillationerasegeneral}]
\end{itemize}

We will evaluate these algorithms on CIFAR-10 [\cite{Krizhevsky09learningmultiple}], Kaggle NeurIPS 2023 Machine Unlearning dataset [\cite{neurips-2023-machine-unlearning}], and MUFAC (Machine Unlearning Facial Attribute Classification) dataset [\cite{choi2023machine}]. 

\section{What will the “contribution” be?}
The main contribution of this project will be a reproducibility report on Machine Unlearning algorithms, specifically Masked Small Gradient with Convolution Transpose (CT), SCRUB, and DELETE, evaluated against the exact unlearning baseline and the 2023 NeurIPS Unlearning Competition winning solution. We also plan to provide a comprehensive survey of these techniques and methods used to evaluate the efficiency of unlearning algorithms. By applying the competition's privacy metrics across diverse datasets, including CIFAR-10, MUFAC, and the Kaggle competition dataset. We will provide a quantitative analysis of the inherent trade-offs each method makes between unlearning efficacy (forgetting), retained model utility (accuracy), and computational efficiency.
\printbibliography
\end{document}
